Relative parabolic factorizations of Chevalley groups

### 2015-05-22 bass-kolster.tex
Written effective bounds for Bass-Kolster

We can estimate the number of elementary root unipotents involved in the decomposition of the previous theorem.
\begin{lemma}
In the assumptions and notation of Theorem~\ref{thm:BassKolster} one can write any element $g$ of 
$\G(\Phi,R,I)$ as a product of an element of $\G(\Delta_1,R,I)$, element $z_{-\alpha}(s, 1)$ for some $s\in I$ and $n(\Phi)$ elementary unipotents $t_{\alpha_i}(s_i)$, where
\begin{itemize}
 \item $n(\rA_\ell)= 4 \ell -1$;
 \item $n(\rC_\ell)= 8 \ell - 5$;
 \item $n(\rD_\ell)= $
\end{itemize}

\cdot \U(\Sigma_1,I)\cdot \U(\Sigma_1^-,I)\cdot\Z \cdot \U(\Sigma^-_1,I)\cdot \U(\Sigma_1, I).\]
\end{lemma}

>>>>

It is easy to see that the proof of the above theorem is effective and gives an estimate of the total number of elementary root unipotents involved in the decomposition.
\begin{cor}
In the assumptions and notation of Theorem~\ref{thm:BassKolster} every element of $\G(\Phi,R,I)$ 
can be factored into a product of one element of $\G(\Delta_1,R,I)$ one element of $Z$ and at most $n(\Phi)$ elementary root unipotents $t_{\alpha}(s)$ of level $I$, where
$n(\rA_\ell)= 4 \ell - 1$, $n(\rB_\ell)= 8 \ell - 5$, $n(\rC_\ell)= 8 \ell - 5$, $n(\rD_\ell)= 8 \ell - 9$. \end{cor}
\begin{proof}
The estimates stated above can be obtain by careful analysis of the proof of the previous theorem.
Moreover, the estimates for $\Phi=\rA_\ell, \rC_\ell$ are immediate.
In the case $\Phi=\rD_\ell$ the proof of Theorem~\ref{thm:BassKolster} implies 
\begin{multline}\nonumber
\G(\Phi,R,I)=  \U(\Sigma_\ell,I) \cdot X_{\alpha_1}(I) \cdot \U(\Sigma_2^-\cap\Delta_1,I) \cdot X_{-\alpha\ssub{\mathrm{max}}}(I) \cdot Z  \cdot \\ \cdot \U(\Sigma_1^-,I) \cdot \U(\Sigma_1,I) \cdot \G(\Delta_1,R,I).
\end{multline}
We can present an element $g$ of $\U(\Sigma_\ell, I)$ as a product of $g_1 \in \U((\Sigma_1 \cup \Sigma_2) \cap \Sigma_\ell)$ and $g_2\in \U(\Sigma_\ell\cap\Delta_1\cap\Delta_2)$.
An examination of the extended Dynkin diagram of $\rD_\ell$ implies that $g_2$ either centralizes or normalizes all factors of the above decomposition (except the last one) and therefore can be safely moved to the right.
On the other hand, $g_1$ is a product of at most $2\ell-3$ elementary unipotents, while the width of $\U(\Sigma_1^\pm, I)$ and $\U(\Sigma_2^-\cap\Delta_1)$ in elementary unipotents does not exceed $2\ell-2$ and $2\ell-4$, respectively.
Summing up these upper bounds we obtain
$$n(\rD_\ell)\leq (2\ell-3) + 1 + (2\ell - 4) + 1 + 2\cdot (2\ell - 2) = 8\ell - 9.$$

The estimate for the case $\Phi=\rB_\ell$ can be obtained by a similar argument. \end{proof}
